---
title: "Flora Cordoleani Data"
author: "Maddee Rubenson (FlowWest)"
date: "`r Sys.Date()`"
output:
  rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE, 
                      message = FALSE)

library(tidyverse)
library(lubridate)
library(knitr)
library(readxl)
library(sf)
library(preyDataProcessing)

options(scipen=999)
```

## Cordoleani Data Standardization

**Datasets provided:**

-   `cage_locations.csv`
-   `cagesalmon_growth.csv`
-   `dailycagefish_data_2019.csv`
-   `dailycagefish_data_2020.csv`
-   `doboall.csv`
-   `tidy_zoop_density.csv`

**Author contact info:**

Flora Cordoleani [[flora.cordoleani\@noaa.gov](flora.cordolean@noaa.gov)]

### Prey Data

Final prey density dataset includes the following variables:

-   `date`: YYYY-MM-DD
-   `gear_type`: the type of gear used to collect zooplankton.
-   `species`: species of zooplankton
-   `prey_density`: density of zooplankton (count/L)
-   `size_class`: size class of zooplankton, determined by mesh size
-   `mesh_size`: mesh size of net used to collect zooplankton
-   `habitat_type`: habitat type of location where zooplankton were collected
-   `lat`: latitude of sampling location
-   `lon`: longitude of sampling location
-   `site`: location description
-   `author`: author of dataset
-   `watershed`: unique watershed name associated with collected data

#### Raw data

```{r}
path <- system.file("extdata", "cordoleani", "tidy_zoop_density.csv", package = "preyDataProcessing")
cordo_prey_raw <- read_csv(path) |> glimpse()

```

#### Standard format

**excluded variables:**

-   removed temperature to include in a standalone dataset that can be mapped to this dataset using `sample_id` and `date`

**notes:**

-   chose to not use interpolated or filled values for prey density

-   Defined `sample_id` as `cage_id`

-   Defined `species` as `taxa`

-   mapped `Canal channel` and `Agriculture` to `agricultural canal`

-   mapped `Channel` and `River Channel` to `perennial instream`

-   mapped `Wetland` to `floodplain`

-   Changed units for `prey_density` from count/m\^3 to count/L

```{r}
cordoleani_zoop <- cordo_prey_raw %>% 
  rename(prey_density = value, # units are organisms/m^3 per Sutter 2020 report 
         sample_id = report_id,
         date = sample_date,
         location = region,
         species = taxa) %>%
  mutate(date = mdy(date),
         prey_density = prey_density/1000, # 1 m^3 = 1000 Liters 
         author = 'Cordoleani',
         habitat_type = ifelse(habitat_type == "Agriculture", 'agricultural canal',
                               ifelse(habitat_type == 'Channel', 'perennial instream',
                                      ifelse(habitat_type == "River Channel", 'perennial instream',
                                             ifelse(habitat_type == 'Wetland', 'floodplain', 'agricultural canal' )))),
         watershed = case_when(location == "Butte Sink" ~ "Butte Creek", 
                            location == "Sacramento River" ~ "Sacramento River", 
                            location == "Feather River" ~ "Feather River", 
                            location %in% c("Lower Bypass", "Upper Bypass") ~ 
                              "Sutter Bypass"), 
         site = paste0(location, "-", sample_id),
         size_class = "meso",
         gear_type = "net throw",
         mesh_size = 150,
         species = tolower(species)) %>% 
         select(-location, -sample_id, -year) 

```

##### Locations Standard Format

**variables removed**

**notes:**

```{r}
path <- system.file("extdata", "cordoleani", "Cage_Locations.csv", package = "preyDataProcessing")
cordoleani_locations_raw <- read_csv(path) 

cordoleani_locations <- cordoleani_locations_raw %>%
  rename(location = Region, 
         sample_id = Report_ID, 
         habitat_type = Type, 
         lon = long) %>%
  select(-Name,  -Report_type) %>%
  mutate(habitat_type = ifelse(habitat_type == "Agriculture", 'agricultural canal',
                               ifelse(habitat_type == 'Channel', 'perennial instream',
                                      ifelse(habitat_type == "River Channel", 'perennial instream',
                                             ifelse(habitat_type == 'Wetland', 'floodplain', 'agricultural canal' ))))) %>%
  mutate(site = paste0(location, "-", sample_id)) %>%
  mutate(site_for_fish = paste0(location, "-", Site)) %>%
  select(-location, -sample_id, -habitat_type) %>%
  glimpse()

```

##### Combine prey data with locations

This produces the final prey density dataset.

```{r}

cordoleani_zoop_final <- cordoleani_zoop |> 
  left_join(cordoleani_locations) |> 
  select(-site_for_fish, -Site)

kable(head(cordoleani_zoop_final, 5))

```

#### QC

**Notes:**

```{r}
summary(cordoleani_zoop_final)
```

#### Data exploration

```{r}
ggplot(cordoleani_zoop_final, aes(x = as.factor(month(date)), y = prey_density)) + 
  geom_point() + 
  facet_grid(~year(date)) + 
  xlab('month') +
  ylab('prey density (count/L)') + 
  ggtitle('Distrubtion of prey density across years collected', 
          subtitle = "data provided by Flora Cordoleani") 

ggplot(cordoleani_zoop_final, aes(x = as.factor(month(date)), y = prey_density)) + 
  geom_point(aes(color = as.factor(year(date)))) + 
  facet_wrap(~habitat_type + year(date)) + 
  xlab('month') +
  ylab('prey density (count/L)') + 
  ggtitle('Distrubtion of prey density across years and habitat types', 
          subtitle = "data provided by Flora Cordoleani") +  
  scale_color_manual('sample year', values=c('darkgreen', 'darkblue', "darkred")) + 
  theme(legend.position = "top")

```

#### Save final dataset

```{r}

save(cordoleani_zoop_final, file = "../../data/cordoleani_prey_data.rda")

cordoleani_prey_data <- cordoleani_zoop_final
usethis::use_data(cordoleani_prey_data, overwrite = TRUE)

```

### Fish Data

[data dictionary - overview of what the data looks like]

#### Raw data

#### Standard format

-   excluded variables:

-   notes:

#### QC

#### Data exploration

[data dictionary - overview of what the data looks like]

#### Raw data

#### Standard format

-   excluded variables:

-   notes:

#### QC

#### Data exploration

### Environmental Data
