---
title: "Prey Data Overview"
author: "Maddee Rubenson (FlowWest)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE, message = FALSE, warning = FALSE
)
library(tidyverse)
library(prey_data_processing)
library(knitr)
```

# Load Data 

## Prey Density 


```{r cars}
monthly_prey_density <- prey_data_processing::monthly_prey_density %>% glimpse

# There are two NAs associated with the month of San Joaquin data because data received was aggregated to year

```

## Fish Data 

```{r}
monthly_fish <- prey_data_processing::monthly_fish_data %>% glimpse

```

## Environmental Data 

```{r}
all_enviro <- prey_data_processing::all_enviro_data %>% glimpse

```

# Explore Variables

## Authors

```{r}

prey_agg_by_author <- prey_data_processing::all_prey_density %>%
  group_by(author) %>%
  summarise_at(vars(prey_density), .funs = c('min', 'max', 'median', 'var'), na.rm = TRUE)

kable(prey_agg_by_author, digits = 1, caption = "summary statistics of prey density values across authors included in the prey data package")

```

## Habitat Type

Habitat types included in the prey data package include: 

* Agricultural canal
* Floodplain
* Perennial instream
* Side channel

Many habitat types were assigned based on the sampling locations provided by authors. 

```{r}

ggplot() +
  geom_boxplot(data = monthly_prey_density, aes(x = as.factor(month), y = median_prey_density), alpha = 0.5) +
  geom_jitter(data = monthly_prey_density, aes(x = as.factor(month), y = median_prey_density, color = habitat_type)) +
  scale_y_continuous(trans='log10') +
  xlab('month') +
  ylab('log(median prey density) (count/L)') +
  scale_color_brewer(palette = 'Dark2') +
  coord_flip() +
  theme_minimal()

prey_data_processing::all_prey_density %>%
  group_by(habitat_type) %>%
  summarise(author = paste0(unique(author), collapse = ", ")) %>% kable()

```

## Watersheds

There are `r length(unique(monthly_prey_density$watershed))` watersheds included in the monthly prey density data, including:

* Butte Creek
* Feather River
* Merced
* North Delta
* Sacramento River
* San Joaquin River
* South Delta
* Stanislaus
* Sutter Bypass
* Yolo Bypass
* Yolo Bypass 2 #TODO: aggregate Yolo and Yolo Bypass 2

```{r}
ggplot() +
  geom_boxplot(data = monthly_prey_density, aes(x = as.factor(month), y = median_prey_density), alpha = 0.5) +
  geom_jitter(data = monthly_prey_density, aes(x = as.factor(month), y = median_prey_density, color = watershed)) +
  scale_y_continuous(trans='log10') +
  ylab('log(median prey density) (count/L)') +
  xlab('month') +
  #facet_wrap(~watershed)+
  scale_color_brewer(palette = 'Dark2')+
  coord_flip()

```

## Temporal Variation 

```{r}
ggplot() +
  geom_boxplot(data = monthly_prey_density, aes(x = as.factor(month), y = median_prey_density), alpha = 0.5) +
  geom_jitter(data = monthly_prey_density, aes(x = as.factor(month), y = median_prey_density, color = year)) +
  scale_colour_continuous(type = 'viridis')+
  scale_y_continuous(trans='log10') +
  ylab('log(median prey density) (count/L)') +
  xlab('month') +
  coord_flip()


```

## Fish Data

```{r}

# mass
ggplot() +
  geom_boxplot(data = monthly_fish, aes(x = as.factor(month), y = mass_median), alpha = 0.5) +
  geom_jitter(data = monthly_fish, aes(x = as.factor(month), y = mass_median, color = habitat_type)) +
  ylab('mass (grams)') +
  xlab('month') +
  scale_color_brewer(palette = 'Dark2')+
  coord_flip()

# fork length (mm)
ggplot() +
  geom_boxplot(data = monthly_fish, aes(x = as.factor(month), y = fork_length_median), alpha = 0.5) +
  geom_jitter(data = monthly_fish, aes(x = as.factor(month), y = fork_length_median, color = habitat_type)) +
  ylab('fork length (mm)') +
  xlab('month') +
  scale_color_brewer(palette = 'Dark2')+
  coord_flip()

```

## Environmental Data 

* Not aggregated by month
* Includes: dissolved oxygen, temperature and flow 

```{r}
# temperature
all_enviro %>% filter(!is.na(temperature)) %>%
  ggplot() +
  geom_line(aes(x = date, y  = temperature)) + 
  facet_wrap(~ author) +
  ylab('temperature (celcius)')

# dissolved oxygen
all_enviro %>% filter(!is.na(do_mg_l)) %>%
  ggplot() +
  geom_line(aes(x = date, y  = do_mg_l)) + 
  facet_wrap(~ author) +
  ylab('dissolved oxygen (mg/l)')

# flow
all_enviro %>% filter(!is.na(flow_cfs)) %>%
  ggplot() +
  geom_line(aes(x = date, y  = flow_cfs)) + 
  facet_wrap(~ author) +
  ylab('flow_cfs')


```
