---
title: "Jason Guignard Data"
author: "Maddee Rubenson (FlowWest)"
date: "`r Sys.Date()`"
output:
  rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE, 
                      message = FALSE)

library(tidyverse)
library(lubridate)
library(knitr)
library(readxl)
library(sf)
library(preyDataProcessing)

options(scipen=999)
```

## Guignard Data Standardization

**Datasets provided:**

-   `drift net sites_to_kml.kml`
-   `Honolulu Bar data.xlsx`

**Author contact info:**

Jason Guignard [jasonguignard\@fishbio.com]

### Prey Data

Final prey density dataset includes the following variables:

-   `date`: YYYY-MM-DD
-   `gear_type`: the type of gear used to collect zooplankton.
-   `species`: species of zooplankton
-   `prey_density`: density of zooplankton (count/L)
-   `size_class`: size class of zooplankton, determined by mesh size
-   `mesh_size`: mesh size of net used to collect zooplankton
-   `habitat_type`: habitat type of location where zooplankton were collected
-   `lat`: latitude of sampling location
-   `lon`: longitude of sampling location
-   `site`: location description
-   `author`: author of dataset
-   `watershed`: unique watershed name associated with collected data

#### Raw data

```{r}

path <- system.file("extdata", 'guignard', "Honolulu Bar data.xlsx", package = "preyDataProcessing")
stanislaus_zoop_raw <- readxl::read_excel(path, sheet = 'Density by location, date,ta_MR')

stanislaus_zoop_raw |> glimpse()

```

#### Standard format

**excluded variables:**

-   `prey_density_total` - chose to remove this and use the `prey_density` disaggregated by species

**notes:**

-   Used `Total Drift Density` column as prey density value which needed to be converted from cubic feet to liters

-   `site` = `location`

```{r}
stanislaus_zoop <- stanislaus_zoop_raw %>%
  rename(location = Location,
         prey_density_total = `Total  Drift Density`,# units are cubic feet
         other_taxa = `Other Taxa`) %>%
  mutate(prey_density_total = prey_density_total/28.3168) %>%  #1ft^3 = 28.3168 Liters  # this is the sum of all species     
  gather(!c('location', 'prey_density_total', 'Date'), key = species, value = value )%>%
  mutate(author = 'Guignard',
         watershed = 'Stanislaus',
         species = tolower(species)) %>%
  rename(prey_density = value,
         date = Date) %>%
  mutate(prey_density = prey_density/0.0283168, 
         site = location,
         gear_type = "net throw",
         mesh_size = 335,
         size_class = "meso/macro") %>% # size class is 335 micrometer mesh. 
  select(-prey_density_total, -location) 
```

##### Locations Standard Format

**variables removed**

**notes:**

-   Match names of locations to the `stanislaus_zoop` table and merge them
-   Manually assign `habitat_type` using best professional judgement

```{r}
path <- system.file("extdata", 'guignard', "drift net sites_to_kml.kml", package = "preyDataProcessing")
locations_raw <- st_read(dsn = path) 
geometry <- sf::st_coordinates(locations_raw) |> 
  cbind(locations_raw) |> 
  rename(lon = X, 
         lat = Y) |> 
  select(-Z, -geometry, -Description) |> 
  sf::st_drop_geometry()

# locaton names of zoop data
unique(stanislaus_zoop$site)

# location names in .kmz 
unique(geometry$Name)

# change the names of the locations dataset to match those in in the zoops table 
geometry[geometry$Name == "H Bar Main Channel", 'Name'] <- 'Main Channel'
geometry[geometry$Name == "Side Channel Upper", 'Name'] <- 'Top Side'
geometry[geometry$Name == "Side Channel Lower", 'Name'] <- 'Lower Side'
geometry[geometry$Name == "Floodplain Upper", 'Name'] <- 'Upper Flood Plain'
geometry[geometry$Name == "Middle Floodplain", 'Name'] <- 'Middle Flood Plain' 
geometry[geometry$Name == "Floodplain Lower", 'Name'] <- 'Lower Flood Plain'

locations <- geometry %>%
  mutate(habitat_type = ifelse(Name %in% c('Main Channel'), 'perennial instream', 
                               ifelse(Name %in% c('Top Side', 'Lower Side'), 'side channel',
                                      ifelse(Name %in% c('Upper Flood Plain', 'Middle Flood Plain', 'Lower Flood Plain'), 'floodplain', NA)))
  ) %>%
  rename(site = Name)




```

##### Combine prey data with locations

This produces the final prey density dataset.

```{r}

guignard_prey_data_final <- stanislaus_zoop |> 
  left_join(locations) 

kable(head(guignard_prey_data_final, 5))

```

#### QC

**Notes:**

-   Data collected in April, May, and June of 2014

```{r}
summary(guignard_prey_data_final)
```

#### Data exploration

```{r}
ggplot(guignard_prey_data_final, aes(x = as.factor(month(date)), y = prey_density)) + 
  geom_point() + 
  facet_grid(~year(date)) + 
  xlab('month') +
  ylab('prey density (count/L)') + 
  ggtitle('Distrubtion of prey density across years collected', 
          subtitle = "data provided by Jason Guignard") 

ggplot(guignard_prey_data_final, aes(x = as.factor(month(date)), y = prey_density)) + 
  geom_point(aes(color = as.factor(year(date)))) + 
  facet_wrap(~habitat_type) + 
  xlab('month') +
  ylab('prey density (count/L)') + 
  ggtitle('Distrubtion of prey density across years and habitat types', 
          subtitle = "data provided by Jason Guignard") +  
  scale_color_manual('sample year', values=c('darkgreen', 'darkblue', "darkred")) + 
  theme(legend.position = "top")

```

#### Save final dataset

```{r}

#save(guignard_prey_data_final, file = "../../data/guignard_prey_data.rda")

guignard_prey_data <- guignard_prey_data_final
usethis::use_data(guignard_prey_data, overwrite = TRUE)

```

### Fish Data

[data dictionary - overview of what the data looks like]

#### Raw data

#### Standard format

-   excluded variables:

-   notes:

#### QC

#### Data exploration

[data dictionary - overview of what the data looks like]

#### Raw data

#### Standard format

-   excluded variables:

-   notes:

#### QC

#### Data exploration

### Environmental Data
