---
output: 
  html_document
---

```{r eval=TRUE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)

library(tidyverse)
library(lubridate)
library(knitr)
library(sf)
library(leaflet)
library(preyDataProcessing)
```

## Habitat Type

### Types:

Sampling locations were assigned the following habitat types:

-   perennial instream

-   floodplain

-   agricultural canal TODO: change to perennial instream?

-   side channel TODO: change to perennial instream?

-   rice field

### Methodology

Habitat types were defined for the following datasets according to the following rulesets.

-   **Cordoleani**: locations (latitude, longitude) were provided by Cordoleani as well as habitat types. The provided habitat types were mapped to a consistent format using the rules:
    -   `Agriculture` = `agricultural canal`

    -   `Channel` = `perennial instream`

    -   `River Channel` = `perennial instream`

    -   `Wetland` = `floodplain`
-   **Corline**:
    -   The locations of the fields (Knaggs Ranch) were provided by Corline

    -   The location of the inlet canal (Toe Drain) was determined using best professional judgement and guidance from Corline et al.

    -   Location of the Sacramento River sampling site was provided in the paper Corline et al. as Sherwood Harbor

    -   The `habitat_type` was defined using best professional judgement from monitoring locations
-   **Zeug**:
    -   The San Joaquin River sampling locations (latitude, longitude) were provided by Zeug

    -   The \`habitat_type\` was defined using best professional judgement from monitoring locations

    -   No locations data was provided for sampling sites on the Merced River
-   **Montgomery**: locations (latitude, longitude) and habitat types were provided by Jacob Montgomery
-   **Guignard**: locations (latitude, longitude) were provided by Guignard. The habitat types were assigned using best professional judgement and following ruleset based off the site location names:
    -   `Main Channel` = `perennial instream`

    -   `Top Side` and `Lower Side` = `side channel` TODO: should we map this as perennial instream?

    -   `Upper Floodplain`, `Middle Floodplain`, `Lower Floodplain` = `floodplain`
-   **Zooper**: locations (latitude, longitude) were provided by the Zooper library. Habitat types were assigned based on the following ruleset:
    -   If the data was collected in Yolo Bypass, it was assigned floodplain

    -   If the data was collected in the North or South Deltas, it was assigned perennial instream

### Map of Data Collection Locations

```{r, include = FALSE}

hab_and_author <- all_prey_density |> 
  select(author, lat, lon, habitat_type) |> 
  distinct() |> 
  mutate(author2 = ifelse(grepl("zooper", author), "Zooper", author), 
         color = case_when(author == "Cordoleani" ~ "#f6eff7", 
                           author == "Corline" ~ "#bdc9e1",
                           author == "Zeug" ~ "#67a9cf", 
                           author == "Montgomery" ~ "#1c9099", 
                           author == "Guignard" ~ "#016c59",
                           grepl("zooper", author) ~ "#12638e"))

```

```{r}
leaflet(data = hab_and_author) %>%
  addProviderTiles(providers$Esri.WorldTopoMap, group = "Map") %>% 
  addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  addCircleMarkers(data = hab_and_author, 
                   color = ~color,
                   lng = ~lon, 
                   lat = ~lat,
                   popup = ~paste0("habitat type: ", habitat_type, "<br>",
                                   "author: ", author)) %>%
   addLegend(colors = ~unique(color), labels = ~unique(author2),
                position = 'topleft', title = 'author')

```

#### Save locations data

```{r}

hab_and_author |> 
  select(-author2, -color) |> 
  write_csv("../data-raw/locations_and_habitat_data.csv")

```
