---
title: "Nicholas Corline Data"
author: "Maddee Rubenson (FlowWest)"
date: "`r Sys.Date()`"
output:  
  rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE, message = FALSE, warning = FALSE
)

library(tidyverse)
library(lubridate)
library(knitr)
library(readxl)
library(sf)
library(preyDataProcessing)

options(scipen=999)
```

## Corline Data Standarization

**Datasets provided:**

-   Zoop2013_2016_NC.xlsx - this is the most up to date dataset used in the prey data aggregation
-   Zoop2013_2016.xlsx
-   Corline2017_Article_ZooplanktonEcologyAndTrophicRe.pdf

**Author contact info:**

Nicholas Corline [njcorline\@vt.edu]

### Prey Data

Final prey density dataset includes the following variables:

-   `date`: YYYY-MM-DD
-   `gear_type`: the type of gear used to collect zooplankton.
-   `species`: species of zooplankton
-   `life_stage`: lifestage of zooplankton species
-   `prey_density`: density of zooplankton (count/L)
-   `size_class`: size class of zooplankton, determined by mesh size
-   `mesh_size`: mesh size of net used to collect zooplankton
-   `habitat_type`: habitat type of location where zooplankton were collected
-   `lat`: latitude of sampling location
-   `lon`: longitude of sampling location
-   `site`: location description
-   `author`: author of dataset
-   `watershed`: unique watershed name associated with collected data

#### Raw data

```{r}
excel_path <- system.file("extdata", "corline", "Zoop2013_2016_NC.xlsx", package = "preyDataProcessing")
corline_raw_zoop <- readxl::read_excel(excel_path) |> glimpse()
```

#### Standard format

**excluded variables:**

-   `Split Fraction`

-   `Class`

-   `Subclass`

-   `Order`

-   `Genus`

-   `Family`

**notes:**

-   Kept `species` column

-   Fixed naming conventions of the sampling locations (`sample_id`)

-   Defined prey density as `Ind_per_L` which was calculated as the ratio of field volume and sample totals

-   Along with `gear_type` included in the dataset are the number of throws (`n_throws`) and mesh size (`mesh_size_microns`). Mesh size is used to determine the `size_class` variable.

```{r}
corline_zoop_tmp <- corline_raw_zoop  |> 
  rename(location = Location, 
         gear_type = Method, 
         life_stage = `Life Stage`) |> 
  mutate(prey_density = Ind_per_L,
         Date = as_date(Date)) |> 
  select(-c(`Split Fraction`:`Volume subsampled (ml)`), -Count, -c(Phylum:Genus)) %>%
  rename(sample_id = Field, 
         species = Species, 
         n_throws = Throws, 
         mesh_size_microns = `Mesh Size (microns)`,
         date = Date) |> 
  select(-`Total Volume (ml)`, -`Ring Size (cm)`, -Ind_per_m3,
         -Ind_per_L, -...25, -...26, -Field_Volume_m3, -Inverse_Fraction, 
         -Sample_Total_Count) |> 
  mutate(author = "Corline",
         size_class = "meso",
         species = tolower(species),
         gear_type = tolower(gear_type),
         mesh_size = mesh_size_microns) 

# Numeric sample Ids must be standardized with Fields. Looks like 2015 data has 1 - 9 as sample Id while data from 2013 has Field 1 - Field 9. This is updated to have the sample_ids as follows, where the NA corresponds to the sites on the Sacramento River. 
tmp_change_names <- corline_zoop_tmp[corline_zoop_tmp$sample_id %in% c(1:10), ] %>%
  mutate(sample_id = paste0('Field ', sample_id))
tmp_names_with_field <- corline_zoop_tmp[!(corline_zoop_tmp$sample_id %in% c(1:10)), ]
corline_zoop_tmp <- bind_rows(tmp_change_names, tmp_names_with_field)

# this fixes the spelling differences in species column 
corline_zoop <- corline_zoop_tmp |> 
  mutate(species = case_when(species %in% c("harpacticoid", "harpacticoida") ~ "harpacticoida",
                             species %in% c("illyocryptidae", "ilyocryptidae") ~ "ilyocryptidae",
                             species %in% c("schaploberis", "schaphloberis") ~ "schaphloberis", 
                             species == "species" ~ NA_character_, 
                             TRUE ~ species))  |> 
  select(-n_throws, -mesh_size_microns) 

```

##### Locations Standard Format

**Notes:**

-   location of the fields (Knaggs Ranch) were provided by author

-   location of inlet canal (Toe Drain) was determined based on best professional judgement and guidance from the paper Corline et al.

-   location of the Sacramento River sampling site was provided in the paper Corline et al. as Sherwood Harbor

-   defined `habitat_type` based on best professional judgement from monitoring locations

```{r}
corline_locations <- data.frame(
  location = c('Knaggs Ranch', 'Sacramento River', 'Toe Drain'),
  habitat_type = c('floodplain', 'perennial instream', 'agricultural canal'), # TODO update these habitat designations per recommendations from J. Montgomery
  lat = c(38.70151, 38.52353, 38.70511),
  lon = c( -121.66996, -121.53840, -121.67041)
)

```

##### Combine prey data with locations

This produces the final prey density dataset.

```{r}

corline_zoop_final <- corline_zoop |> 
  left_join(corline_locations) |>  
   mutate(watershed = ifelse(location == "Sacramento River", "Sacramento River", "Yolo Bypass"),
         site = ifelse(location == "Sacramento River", "Sherwood Harbor", location)) %>%
  mutate(site = paste0(location, "-", sample_id)) %>%
  select(-sample_id, -location) 

kable(head(corline_zoop_final, 5))

```

#### QC

**Notes:**

-   all prey data was collected January through April

```{r}
summary(corline_zoop_final)
```

#### Data exploration

```{r}
ggplot(corline_zoop_final, aes(x = month(date), y = prey_density)) + 
  geom_point() + 
  facet_grid(~year(date)) + 
  xlab('month') +
  ylab('prey density (count/L)') + 
  ggtitle('Distrubtion of prey density across years collected', 
          subtitle = "data provided by Nicholas Corline") 

ggplot(corline_zoop_final, aes(x = month(date), y = prey_density)) + 
  geom_point() + 
  facet_grid(~habitat_type + year(date)) + 
  xlab('month') +
  ylab('prey density (count/L)') + 
  ggtitle('Distrubtion of prey density across years and habitat types', 
          subtitle = "data provided by Nicholas Corline") 

```

#### Save final dataset

```{r}

save(corline_zoop_final, file = "../../data/corline_prey_data.rda")

corline_prey_data <- corline_zoop_final
usethis::use_data(corline_prey_data, overwrite = TRUE)

```

### Fish Data

[data dictionary - overview of what the data looks like]

#### Raw data

#### Standard format

-   excluded variables:

-   notes:

#### QC

#### Data exploration

[data dictionary - overview of what the data looks like]

#### Raw data

#### Standard format

-   excluded variables:

-   notes:

#### QC

#### Data exploration

### Environmental Data
