---
title: "Steve Zeug Data"
author: "Maddee Rubenson (FlowWest)"
date: "`r Sys.Date()`"
output: 
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE, 
                      message = FALSE)

library(tidyverse)
library(lubridate)
library(knitr)
library(readxl)
library(sf)
library(preyDataProcessing)

options(scipen=999)
```

## Zeug Data Standardization

**Datasets provided:**

-   Merced River:
    -   `Flow_meter_Drift.xlsx`

    -   `Merced invert data_environment.xls`

    -   `salmon data_final(Josie).xlsx`

    -   `Temperature.xlsx`
-   San Joaquin River
    -   `summary_zooplankton_10_19_17.xlsx`

**Author contact info:**

Steve Zeug [stevez\@fishsciences.net]

## Prey Data

### San Joaquin River Dataset:

Final prey density for the San Joaquin River dataset provided by Zeug includes the following variables:

-   `year`: calendar year data was collected

-   `week`: week data was collected

-   `gear_type`: the type of gear used to collect zooplankton.

-   `species`: species of zooplankton

-   `prey_density`: density of zooplankton (count/L)

-   `size_class`: size class of zooplankton, determined by mesh size

-   `mesh_size`: mesh size of net used to collect zooplankton

-   `habitat_type`: habitat type of location where zooplankton were collected

-   `lat`: latitude of sampling location

-   `lon`: longitude of sampling location

-   `site`: location description

-   `author`: author of dataset

-   `watershed`: unique watershed name associated with collected data

#### Raw data

```{r}
path <- system.file("extdata", "zeug", "san_joaquin", "summary_zooplankton_10_19_17.xlsx", package = "preyDataProcessing")
zeug_prey_data_raw <- readxl::read_excel(path, sheet = "CountLiter2016", range = "A1:H56") |> glimpse()

```

#### Standard format

**excluded variables:**

**notes:**

-   column `CountPerVolume` was used for `prey_density` and determined to have units of total prey/liter and kept as such

-   since we do not have sampling dates, I kept in `Year` and `Week`. We might want to remove these for the full standardized dataset.

-   added lat/longs provided by Steve Zeug

-   manually assigned `habitat_types` based on best professional judgement (see map below for verification)

```{r}
zeug_prey_data_san_joaquin <- zeug_prey_data_raw %>%
  rename(sample_id = Reach, 
         species = TaxGrp_1,
         count = TotalTaxbyReach,
         prey_density = CountperVolume) %>%# currently total prey/l
  select(sample_id, species, prey_density, Year, Week) %>% 
  mutate(author = "Zeug",
         watershed = 'San Joaquin',
         site = sample_id,
         size_class = "micro",
         species = tolower(species),
         gear_type = "10 liter Schindler-Patalas",
         mesh_size = 64) %>%
  select(-sample_id) |> 
  rename(week = Week, 
         year = Year)

```

##### Locations Standard Format

**variables removed**

**notes:**

```{r}
sj_locations <- data.frame(
  site = c('1A', '2A', '4B-CS', '4B-ESB'),
  habitat_type = c('perennial instream', 'perennial instream', 'side channel',  'perennial instream'),
  lat = c(36.860461, 36.798219, 37.166257, 37.166817),
  lon = c(-119.847043, -120.16085, -120.619793, -120.631145)
)
```

##### Combine prey data with locations

This produces the final prey density dataset.

```{r}

zeug_prey_data_san_joaquin_final <- zeug_prey_data_san_joaquin |> 
  left_join(sj_locations) 

kable(head(zeug_prey_data_san_joaquin_final, 5))

```

#### QC

**Notes:**

```{r}
summary(zeug_prey_data_san_joaquin_final)
```

#### Data exploration

```{r}
ggplot(zeug_prey_data_san_joaquin_final, aes(x = as.factor(week), y = prey_density)) + 
  geom_point(aes(color = habitat_type)) + 
  facet_grid(~year) + 
  xlab('month') +
  ylab('prey density (count/L)') + 
  ggtitle('Distrubtion of prey density across years and habitat types collected', 
          subtitle = "data provided by Steve Zeug for the San Joaquin River") +
  scale_color_manual('habitat type', values=c('darkred', 'darkblue')) + 
  theme(legend.position = "top")

```

#### Save final dataset

```{r}

save(zeug_prey_data_san_joaquin_final, file = "../../data/zeug_sanJoaquin_prey_data.rda")

```

### Merced River Dataset:

Final prey density for the Merced River dataset provided by Zeug includes the following variables:

-   `date` : date data was collected

-   `gear_type`: the type of gear used to collect zooplankton.

-   `prey_density`: density of zooplankton (count/L)

-   `size_class`: size class of zooplankton, determined by mesh size

-   `mesh_size`: mesh size of net used to collect zooplankton

-   `habitat_type`: habitat type of location where zooplankton were collected

-   `site`: location description

-   `author`: author of dataset

-   `watershed`: unique watershed name associated with collected data

#### Raw data

```{r}
path <- system.file("extdata", "zeug", "merced", "Merced invert data_environment.xls", package = "preyDataProcessing")
zeug_merced_prey_data_raw <- readxl::read_excel(path, .name_repair = 'minimal', sheet = "erin_modified",  col_names = FALSE)

dates <- janitor::excel_numeric_to_date(as.numeric(zeug_merced_prey_data_raw[1,3:ncol(zeug_merced_prey_data_raw)]), date_system = 'mac pre-2011') %>% as.character()

sites <- zeug_merced_prey_data_raw[2, 3:ncol(zeug_merced_prey_data_raw)]

col_names <- append(
  c('Species', 'Measurement'),
  paste(dates, sites, sep = '__')
)

names(zeug_merced_prey_data_raw) <- col_names

# view subset of data
zeug_merced_prey_data_raw |> glimpse()

```

#### Standard format

**excluded variables:**

**notes:**

-   `gear_type` was extracted from column name

-   TODO: check units for prey density. Currently using No. m-2 (subsam) and dividing by provided volume (m\^3) and converting to count/L.

-   TODO: do not have location data for Merced sites

```{r}
zeug_merced_prey_data <- zeug_merced_prey_data_raw[3:nrow(zeug_merced_prey_data_raw),] %>%
  fill(Species, .direction = 'down') %>% #TODO make sure species is accurate(looks like fill is tricky with xlsx format)
  filter(Measurement == 'No. m-2 (subsam)') %>% # is this the value we want for prey density??
  pivot_longer(cols = !c('Species', 'Measurement'), names_to = "date__site") %>%
  separate(col = date__site, into = c('date', 'site'), sep = '__') %>%
  rename(prey_density = value, #TODO: what is this unit, No. m-2 (subsam) ??
        date = date,
         sample_id = site,
         species = Species) %>%
  mutate(author = 'Zeug',
         life_stage = ifelse(grepl('Adult', species), 'Adult',
                             ifelse(grepl('Larvae', species), 'Larvae',
                                    ifelse(grepl('Pupae', species), 'Pupae',
                                           ifelse(grepl('Nymph', species), 'Nymph', NA)))),
         gear_type = ifelse(grepl('Drift', sample_id), 'Drift',
                            ifelse(grepl('Benthic', sample_id), 'Benthic', NA))) %>%
  mutate(tmp_name = gsub('Drift', '', sample_id), # extract sample_id
         name = stringr::str_split(tmp_name, " ") %>% map_chr(., 1),
         sample_id = ifelse(name == '', tmp_name, name ),
         size_class = "macro",
         watershed = "Merced",
         gear_type = tolower(gear_type),
         life_stage = tolower(life_stage),
         species = tolower(species),
         habitat_type = "perennial instream") %>%
  select(-c(tmp_name, name)) %>%
  mutate(sample_id = gdata::trim(sample_id),
         mesh_size = 500) %>%
  filter(gear_type == "drift") %>%
  mutate(gear_type = "net throw") 

# Load in volumes sampled from separate xlsx sheet and use it to calculated count/L 
path <- system.file("extdata", "zeug", "merced", "Flow_meter_Drift.xlsx", package = "preyDataProcessing")
vols <- readxl::read_excel(path, skip = 4) %>%
  janitor::clean_names()

vols <- vols %>% select(site, corrected_voume_m3_s_1) %>%
  filter(!is.na(corrected_voume_m3_s_1)) |> 
  rename(sample_id = site)

zeug_merced_prey_data_final <- zeug_merced_prey_data |> 
  left_join(vols) |> 
  mutate(prey_density = as.numeric(prey_density)/corrected_voume_m3_s_1/1000, # convert from m^3 to L
         date = as.Date(date)) %>%
  rename(site = sample_id) %>%
  select(-corrected_voume_m3_s_1, -Measurement)

kable(head(zeug_merced_prey_data_final, 5))


```

##### Locations Standard Format

There is no location data available for the Merced data provided by Zeug.

#### QC

**Notes:**

-   All data was collected on 2009-04-04

```{r}
summary(zeug_merced_prey_data_final)
```

#### Data exploration

```{r}
ggplot(zeug_merced_prey_data_final, aes(x = as.factor((date)), y = prey_density)) + 
  geom_point(aes(color = habitat_type)) + 
  facet_grid(~year(date)) + 
  xlab('month') +
  ylab('prey density (count/L)') + 
  ggtitle('Distrubtion of prey density across years and habitat types collected', 
          subtitle = "data provided by Steve Zeug for the Merced River") +
  scale_color_manual('habitat type', values=c('darkred', 'darkblue')) + 
  theme(legend.position = "top")

```

#### Save final dataset

```{r}

save(zeug_merced_prey_data_final, file = "../../data/zeug_merced_prey_data.rda")

```

### Fish Data

[data dictionary - overview of what the data looks like]

#### Raw data

#### Standard format

-   excluded variables:

-   notes:

#### QC

#### Data exploration

[data dictionary - overview of what the data looks like]

#### Raw data

#### Standard format

-   excluded variables:

-   notes:

#### QC

#### Data exploration

### Environmental Data
