---
title: "Zooper Data"
author: "Maddee Rubenson (FlowWest)"
date: "`r Sys.Date()`"
output:
  rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE, 
                      message = FALSE)

library(tidyverse)
library(lubridate)
library(knitr)
library(readxl)
library(sf)
library(preyDataProcessing)

options(scipen=999)
```

## Zooper Data Standardization

**Datasets provided:**

-   Data downloaded from Zooper library [[Intergency Ecological Program](https://github.com/InteragencyEcologicalProgram/zooper)]

**Author contact info:**

-   NA

### Prey Data

Final prey density dataset includes the following variables:

-   `date`: YYYY-MM-DD
-   `species`: species of zooplankton
-   `prey_density`: density of zooplankton (count/L)
-   `size_class`: size class of zooplankton, determined by mesh size
-   `habitat_type`: habitat type of location where zooplankton were collected
-   `lat`: latitude of sampling location
-   `lon`: longitude of sampling location
-   `site`: location description
-   `author`: author of dataset
-   `watershed`: unique watershed name associated with collected data

#### Raw data

```{r}

# run the Zooper libary query: 
# zoop_data <- Zoopsynther(Data_type = "Community", 
#                        Sources = c("EMP", "FRP", 
#                                    "FMWT", "STN", "20mm"), 
#                        Size_class = c("Micro", "Meso", "Macro"), 
#                        Date_range = c("1990-10-01", "2020-09-30"))

# save the zooper results: 
#saveRDS(zoop_data, 'zoop_data.rds')

# read pre-saved rds file back in:
path <- system.file("extdata", "zooper", "zoop_data.rds", package = "preyDataProcessing")
zoop_data <- readRDS(path) 

zoop_data |>  glimpse()

```

#### Standard format

**excluded variables:**

-   removed environmental variables

-   `CPUE`

-   `Phylum`

-   `Class`

-   `Order`

-   `Family`

-   `Genus`

-   `Volume`

-   `Undersampled`

**notes:**

-   convert from count/m\^3 to count/L
-   extract `temperature` into unique environmental dataset
-   `Taxname`= `species`

```{r}
zooper <- zoop_data %>%
  select(date = Date, prey_per_unit_volume = CPUE, species = Taxname, life_stage = Lifestage,
         size_class = SizeClass, author = Source, site = SampleID, 
         temperature = Temperature, lat = Latitude, lon = Longitude) %>% 
  mutate(author = paste0('zooper: ', author), 
         species = tolower(species), 
         size_class = tolower(size_class), 
         life_stage = tolower(life_stage), 
         prey_density = prey_per_unit_volume/1000)  # convert to count/L from count/m^3
```

##### Locations Standard Format

**variables removed**

**notes:**

-   Yolo Bypass 2 = floodplain

-   North and South Deltas = perennial instream

-   Data is clipped to delta extents

```{r}
path <- system.file("extdata", "zooper", "shape_files", "delta_extents", "HabitatExtentsForCorreigh.shp", package = "preyDataProcessing")
delta_extents <- rgdal::readOGR(path) %>%
  st_as_sf(coords = c("longitude", "latitude"), dim = "XY", crs = 4326) %>%
  st_transform("WGS84")

sf_zooper_points <- st_as_sf(zooper %>% filter(!is.na(lon), !is.na(lat)), coords = c('lon', 'lat'), crs = 4326, remove = FALSE)

zooper_with_watershed <- st_join(delta_extents, sf_zooper_points) %>%
  st_set_geometry(NULL) %>%
  glimpse()


```

##### Combine prey data with locations

This produces the final prey density dataset.

```{r}

zooper_prey_data <- zooper_with_watershed %>%
    select(-Id, -prey_per_unit_volume, -temperature) %>%
    mutate(habitat_type = ifelse(watershed == 'Yolo Bypass 2', 'floodplain', 'perennial instream')) %>%
    filter(!is.na(author)) # filter out watershed designations with no intersection

kable(head(zooper_prey_data, 5))

```

#### QC

**Notes:**

-   Data collected between 1990 and 2020

```{r}
summary(zooper_prey_data)
```

#### Data exploration

```{r}
ggplot(zooper_prey_data, aes(x = as.factor(month(date)), y = prey_density)) + 
  geom_point() + 
  facet_wrap(~year(date)) + 
  xlab('month') +
  ylab('prey density (count/L)') + 
  ggtitle('Distrubtion of prey density across years collected', 
          subtitle = "data provided by Zooper Library") 

ggplot(zooper_prey_data, aes(x = as.factor(month(date)), y = prey_density)) +
  geom_point() +
  facet_wrap(~habitat_type) +
  xlab('month') +
  ylab('prey density (count/L)') +
  ggtitle('Distrubtion of prey density across habitat types',
          subtitle = "data provided by Zooper Library") +
  theme(legend.position = "top")

```

#### Save final dataset

```{r}

save(zooper_prey_data, file = "../../data/zooper_prey_data.rda")

```

### Fish Data

[data dictionary - overview of what the data looks like]

#### Raw data

#### Standard format

-   excluded variables:

-   notes:

#### QC

#### Data exploration

[data dictionary - overview of what the data looks like]

#### Raw data

#### Standard format

-   excluded variables:

-   notes:

#### QC

#### Data exploration

### Environmental Data
