---
title: "Zooper Data"
author: "Erin Cain"
date: '2022-05-02'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(zooper)
```

## Zooper Data

The following Rmarkdown explores adding [Zooper data](https://github.com/InteragencyEcologicalProgram/zooper) into preyData package.

We pull the following datasets using the `zoopsynther()` function

```{r}
zoop_data <- Zoopsynther(Data_type = "Community", 
                       Sources = c("EMP", "FRP", 
                                   "FMWT", "STN", "20mm"), 
                       Size_class = c("Micro", "Meso", "Macro"), 
                       Date_range = c("1990-10-01", "2020-09-30"))
```

### Prey Data

CPUE = count/m\^3 (this is consistent with out prey_per_unit_volume)

```{r}
selected_zoop <- zoop_data %>% 
  select(date = Date, prey_per_unit_volume = CPUE, taxon = Taxname, life_stage = Lifestage, size_class = SizeClass, source = Source, sample_id = SampleID) %>% 
  glimpse
```

### Environmental Data: 

We selected environmental data from the full zooper dataset. Environmental data can be joined to prey data using the `sample_id` column.

```{r}
selected_environmental <- zoop_data %>% 
  select(date = Date, sample_id = SampleID, 
         disolved_oxygen = DO, temperature = Temperature, 
         salinity = SalSurf) %>% glimpse
```

### Map Sites: 

Questions: What sections are we interested in keeping? We can then assign to the corresponding DSM location if applicable.

```{r}
selected_sites <- zoop_data %>% 
  select(Source, Station, Latitude, Longitude) %>% 
  mutate(color = case_when(Source == "EMP" ~ "#f6eff7", 
                           Source == "FRP" ~ "#bdc9e1",
                           Source == "FMWT" ~ "#67a9cf", 
                           Source == "STN" ~ "#1c9099", 
                           Source == "20mm" ~ "#016c59")) %>%
  distinct()

library(leaflet)
leaflet() %>%
  addTiles() %>%
  addCircleMarkers(data = selected_sites, 
             color = ~color,
             lng = ~Longitude, 
             lat = ~Latitude,
             popup = ~Station)
```

## Save Data Objects as Rds

```{r, eval = FALSE}
write_rds(selected_zoop, "../data/zooper_prey.rds")
write_rds(selected_environmental, "../data/zooper_environmental.rds")
```
