---
title: "Jacob Montgomery Data"
author: "Maddee Rubenson (FlowWest)"
date: "`r Sys.Date()`"
output:
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE, 
                      message = FALSE)

library(tidyverse)
library(lubridate)
library(knitr)
library(readxl)
library(sf)
library(preyDataProcessing)

options(scipen=999)

theme_set(theme_minimal())
```

## Montgomery Data Standardization

**Datasets provided:**

-   [Fish Food on Floodplain Farm Fields 2019 and 2021](https://portal.edirepository.org/nis/mapbrowse?packageid=edi.996.2)

**Author contact info:**

Jacob Montgomery [[jacob\@caltrout.org](jacob@caltrout.org)]

### Prey Data

Final prey density dataset includes the following variables:

-   `date`: YYYY-MM-DD
-   `gear_type`: the type of gear used to collect zooplankton.
-   `species`: species of zooplankton
-   `life_stage`: lifestage of zooplankton species
-   `prey_density`: density of zooplankton (count/L)
-   `size_class`: size class of zooplankton, determined by mesh size
-   `mesh_size`: mesh size of net used to collect zooplankton
-   `habitat_type`: habitat type of location where zooplankton were collected
-   `lat`: latitude of sampling location
-   `lon`: longitude of sampling location
-   `site`: location description
-   `author`: author of dataset

#### Raw data

```{r}
# this a script that sources the EDI data
source('data-raw//montgomery/food_4_fish_data_access.R')

montgomery_prey_data |> glimpse()
```

#### Standard format

**excluded variables:**

-   removed environmental variables like salinity, turbidity
-   `start_rotation`
-   `end_rotation`
-   `notes`
-   `zoop_score_1_10`
-   `time`

**notes:**

-   converted from count/m\^3 to count/L
-   extracted `temperature` and `dissolved oxygen` from prey density dataset but will include them in environmental dataset

```{r}
montgomery_prey_data_process <- montgomery_prey_data %>%
  separate(id, c('location', 'Date'), '_') %>% 
  mutate(date = mdy(Date)) %>% 
  gather(!c('location', 'Date', 'date':'volume_sampled'), key = species, value = value) %>%
  separate(species, c('species', 'life_stage'), '_') %>%
  select(-c(Date, ec_s_cm:do_sat, p_h:volume_sampled)) %>%
  rename(temperature = temp_c,
         prey_density = value) %>%
  mutate(author = "Montgomery", 
         site = location,
         size_class = "meso",
         gear_type = "net throw",
         mesh_size = 150,
        prey_density = prey_density/1000) |> #1 m^3 = 1000 L 
  select(date, site, species:author, size_class, mesh_size, gear_type) 

```

##### Locations Standard Format

**variables removed**

-   `purpose`

-   `habitat_type`: this is the habitat type originally defined by Montgomery but redefined to fit this project needs

**notes:**

```{r}

locations <- readxl::read_excel('data-raw/montgomery/F4F2021_LocationLookupTable_20221108.xlsx') |> janitor::clean_names() |> 
  separate(lat_lon_utm, sep = ", ", c("lat", "lon")) |> 
  select(-x3, -purpose, -habitat_type) |> 
  rename(habitat_type = habitat_type_2,
         site = location) |> 
  mutate(habitat_type = tolower(habitat_type),
         lat = as.numeric(lat),
         lon = as.numeric(lon)) |> glimpse()
```

##### Combine prey data with locations

This produces the final prey density dataset.

```{r}

montgomery_prey_data_final <- montgomery_prey_data_process |> 
  left_join(locations) |> 
  filter(!is.na(habitat_type)) |> # filters out site MNT3ASource which has a prey_density of 0 and no site location or habitat type
  filter(prey_density > 0) 

kable(head(montgomery_prey_data_final, 5))

```

#### QC

**Notes:**

```{r}
summary(montgomery_prey_data_final)
```

#### Data exploration

##### All prey density data

```{r}
ggplot(montgomery_prey_data_final, aes(x = as.factor(month(date)), y = prey_density)) + 
  geom_point(alpha = 0.4) + 
  facet_grid(~year(date)) + 
  xlab('month') +
  ylab('prey density (count/L)') + 
  ggtitle('Distrubtion of prey density across years collected', 
          subtitle = "data provided by Jacob Montgomery") 


ggplot(montgomery_prey_data_final, aes(x = as.factor(month(date)), y = prey_density)) + 
  geom_point(aes(color = as.factor(year(date))), alpha = 0.4) + 
  facet_wrap(~habitat_type) + 
  xlab('month') +
  ylab('prey density (count/L)') + 
  ggtitle('Distrubtion of prey density across years and habitat types', 
          subtitle = "data provided by Jacob Montgomery") +  
  scale_color_manual('sample year', values=c('darkgreen', 'darkblue', "darkred", "orange")) + 
  theme(legend.position = "top")

```

##### Subset of prey density data with outliers removed

```{r}
montgomery_prey_data_final |> 
  filter(prey_density <= 0.22635 & prey_density > 0) |> 
ggplot(aes(x = as.factor(month(date)), y = prey_density)) + 
  geom_boxplot(alpha = 0.4) + 
  facet_grid(~year(date)) + 
  xlab('month') +
  ylab('prey density (count/L)') + 
  ggtitle('Distribution of prey density across years collected - outliers removed', 
          subtitle = "data provided by Jacob Montgomery") 


montgomery_prey_data_final |> 
  filter(prey_density <= 0.22635 & prey_density > 0) |> 
ggplot(aes(x = as.factor(month(date)), y = prey_density)) + 
  geom_boxplot() + 
  facet_wrap(~habitat_type) + 
  xlab('month') +
  ylab('prey density (count/L)') + 
  ggtitle('Distribution of prey density across years and habitat types - outliers removed', 
          subtitle = "data provided by Jacob Montgomery") +  
  theme(legend.position = "top")

```

#### Save final dataset

```{r}

# save(montgomery_prey_data_final, file = "../../data/montgomery_prey_data.rda")

montgomery_prey_data <- montgomery_prey_data_final
usethis::use_data(montgomery_prey_data, overwrite = TRUE)

```

## Upcoming Updates

* include fish weight metrics
* include environmental datasets
